---
output: github_document
  md_document:
    variant: 
---

# Purpose: for the exam

Purpose of this work folder.

Ideally store a minimum working example data set in data folder.

Add binary files in bin, and closed R functions in code. Human Readable settings files (e.g. csv) should be placed in settings/


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(tidyverse)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))


Texevier::create_template(directory = "/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/exam prep",
                          template_name = "Project_Question1",
                          build_project = T, open_project = T)

Texevier::create_template(directory = "/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/exam prep",
                          template_name = "Project_Question2",
                          build_project = T, open_project = T)

Texevier::create_template(directory = "/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/exam prep",
                          template_name = "Project_Question3",
                          build_project = T, open_project = T)

Texevier::create_template(directory = "/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/exam prep",
                          template_name = "Project_Question4",
                          build_project = T, open_project = T)

```

# Question 1
## Prep data
```{r}
library(patchwork)
library(purrr)
library(ggplot2)
library(dplyr)
library(pacman)
library(ggpubr)
library(tseries)  
library(dynlm)
library(zoo)
library(sandwich)
library(car)
library(vars)
library(dplyr)
library(gridExtra)
library(grid)
library(extrafont)
library(urca)
library(lmtest)
library(plm)
library(lfe)
library(geepack)

death_cause <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/Covid/Deaths_by_cause.csv")

covid_data <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/Covid/owid-covid-data.csv")

covid_description <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/Covid/covid_data_description.csv")

View(covid_data)

```

Part A: See if you can provide some insights using the dataset into how African countries’ experience
differed from other regions.

```{r}




#now put it into a function
plot_covid_continent_data <- function(covid_data) {
    continents=c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")
  continent_data <- covid_data %>%
    filter(continent %in% continents) %>%
    group_by(continent) %>%
    summarise(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
              new_cases = sum(new_cases_smoothed_per_million, na.rm = TRUE),
              avg_stringency = mean(stringency_index, na.rm = TRUE))

  ggplot(continent_data) +
    geom_point(aes(x = new_deaths, y = new_cases, size = avg_stringency, color = continents)) +
    labs(x = "New Deaths", y = "New Cases", size = "Average Stringency", color = "Continent") +
    theme_minimal() +
    scale_x_continuous(labels = function(x) format(x, scientific = FALSE)) +
    scale_y_continuous(labels = function(y) format(y, scientific = FALSE))
}

# Usage
plot_covid_continent_data(covid_data)


#looking in more detail regarding vaccination and gdp per capita

library(ggplot2)

analyze_covid_data <- function(covid_data) {
  # Filter relevant columns
    continents=c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")
 

  # Calculate summary statistics
  summary_stats <- covid_data %>%
    filter(continent %in% continents) %>%
    group_by(continent) %>%
    summarize(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
              new_cases = sum(new_cases_smoothed_per_million, na.rm = TRUE),
              vaccination_rollout = sum(new_vaccinations_smoothed_per_million, na.rm = TRUE),
              avg_gdp_per_capita = sum(gdp_per_capita, na.rm = TRUE))

  # Create the plot
  plot <- ggplot(summary_stats) +
    geom_point(aes(x = avg_gdp_per_capita, y = vaccination_rollout, size = new_deaths, color = continent)) +
    labs(x = "GDP per capita", y = "Vaccination Rollout", size = "New Deaths", color = "Continent") +
    theme_minimal() +
    scale_x_continuous(labels = function(x) format(x, scientific = FALSE)) +
    scale_y_continuous(labels = function(y) format(y, scientific = FALSE))

  # Analyze the data
  analysis <- paste("The analysis of COVID-19 data reveals the following:\n",
                    "1. The average number of new deaths ranges from",
                    min(summary_stats$new_deaths, na.rm = TRUE),
                    "to", max(summary_stats$new_deaths, na.rm = TRUE),
                    "across different continents.\n",
                    "2. The average number of new cases ranges from",
                    min(summary_stats$new_cases, na.rm = TRUE),
                    "to", max(summary_stats$new_cases, na.rm = TRUE),
                    "across different continents.\n",
                    "3. The total number of vaccinations ranges from",
                    min(summary_stats$vaccination_rollout, na.rm = TRUE),
                    "to", max(summary_stats$vaccination_rollout, na.rm = TRUE),
                    "across different continents.\n",
                    "4. The average GDP per capita ranges from",
                    min(summary_stats$avg_gdp_per_capita, na.rm = TRUE),
                    "to", max(summary_stats$avg_gdp_per_capita, na.rm = TRUE),
                    "across different continents.")

  # Print the analysis
  cat(analysis, "\n")

  # Return the plot
  return(plot)
}

analyze_covid_data(covid_data)



```
##Part B
Provide some insights into whether countries with specific concentrated groupings (e.g. more
poverty, higher prevalence of smokers, higher general life expectancy and elderly populations)
displayed distinct patterns in the severity of their Covid experience.

```{r}
#testing for one example
#figuring out if population density will work

library(ggplot2)

library(dplyr)

library(dplyr)

# Filter dataset to include only relevant columns
library(dplyr)

Plot_Cases_Density <- function(covid_data) {
    continents <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")
  
  df <- covid_data %>%
    filter(continent %in% continents) %>%
    group_by(continent, location) %>%
    summarise(total_cases = max(new_cases_smoothed_per_million, na.rm = TRUE),
              average_density = mean(population_density, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(continent) %>%
    mutate(mean_density = mean(average_density, na.rm = TRUE))
  
  p <- ggplot(df, aes(x = continent, y = total_cases, fill = continent)) +
      geom_jitter()+
    geom_boxplot() +
    geom_text(aes(y = max(mean_density), label = round(mean_density)),
              color = "black", vjust = -1.5, size = 3) +
    labs(title = "Boxplot of COVID-19 Cases by Continent",
         x = "Continent",
         y = "COVID-19 Cases") +
    theme_bw() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 12),
          axis.title = element_text(size = 12, face = "bold"))+
      scale_y_continuous(labels = function(y) format(y, scientific = FALSE))+
    coord_cartesian(ylim = c(0, 10000))
  
  print(p)
}


Plot_Cases_Density(covid_data)

Extract_Greatest_New_Cases <- function(covid_data) {
  aver_dens <- covid_data %>%
    group_by(location, continent) %>%
    summarise(average_density = mean(population_density, na.rm = TRUE),
              total_cases = mean(new_cases_smoothed_per_million, na.rm = TRUE)) %>%
    filter(!is.na(average_density)) %>% 
    mutate(continent = first(continent)) %>%
    ungroup() %>%
    group_by(continent) %>%
    mutate(greatest_new_cases = max(total_cases, na.rm = TRUE)) %>%
    filter(total_cases == greatest_new_cases) %>%
    dplyr::select(continent, location, greatest_new_cases, average_density) %>%
    ungroup()

  return(aver_dens)
}


# Usage example
extracted_data <- Extract_Greatest_New_Cases(covid_data)

#Now plot regression

library(cowplot)
Regression_Plot <- function(covid_data) {
  # Filter out missing or infinite values
  filtered_data <- covid_data %>%
    filter(!is.na(new_cases_smoothed_per_million), !is.na(population_density))
  
  # Perform the regression for each continent
  regression_models <- filtered_data %>%
    group_by(continent) %>%
    nest() %>%
    mutate(regression = map(data, ~lm(new_cases_smoothed_per_million ~ population_density, data = .x)))
  
  # Plot the regression lines and scatter plots for each continent
  plots <- regression_models %>%
    mutate(
      plot = map2(data, regression, ~{
        ggplot(.x, aes(x = population_density, y = new_cases_smoothed_per_million)) +
          geom_point() +
          geom_smooth(method = "lm", se = FALSE, color = "blue") +
          labs(x = "Population Density", y = "Total Cases", title = paste("Regression: Total Cases vs Population Density -", .y$continent)) +
          theme_bw()
      })
    ) %>%
    pull(plot)
  
  # Combine and arrange the plots
  plot_grid(plotlist = plots, ncol = 2)
}

# Call the function with your COVID-19 data
Regression_Plot(covid_data)





try <- function(covid_data) {
  conts <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")
  correlations <- vector("numeric", length(conts))
  
  for (i in seq_along(conts)) {
    filt_data <- covid_data %>%
      filter(continent == conts[i], !is.na(population_density), !is.na(new_cases_smoothed_per_million))
     
    
    correlations[i] <- cor(filt_data$population_density, filt_data$new_cases_smoothed_per_million, method = "spearman")
  }
  
  result <- data.frame(Continent = conts, correlation = correlations)
  return(result)
}

result <- try(covid_data)
print(result)



library(ggplot2)
library(dplyr)

continents <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")

scatterplot <- covid_data %>%
  filter(continent %in% continents, !is.na(population_density), !is.na(new_cases_smoothed_per_million)) %>% 
  mutate(correlation = round(cor(population_density, new_cases_smoothed_per_million, method = "spearman", use = "complete.obs"), 4)) %>%
  mutate(relationship = ifelse(correlation < 0.5 && correlation > 0, "Weak positive",
                              ifelse(correlation >= 0.5 && correlation <= 1, "Strong positive",
                                     ifelse(correlation > -0.5 && correlation < 0, "Weak negative",
                                            ifelse(correlation <= -0.5 && correlation >= -1, "Strong negative",
                                                   "No correlation"))))) 

scatterplot <- ggplot(scatterplot, aes(x = population_density, y = new_cases_smoothed_per_million)) +
  geom_point() +
  geom_label(aes(label = paste0("Correlation: ", correlation)), 
             size = 6, alpha = 0.8, fill = "red") +
  geom_label(aes(label = paste0("Strength of relationship: ", relationship)), 
             size = 6, alpha = 0.8, fill = "red") +
  theme_bw() +
  labs(title = "Relationship between New Cases and Population Density",
       x = "Population Density",
       y = "New Covid-19 Cases")

scatterplot




#I realised that population density can me a misleading measure, instead use other measures
severity <- covid_data %>%
  filter(!is.na(excess_mortality) & !is.na(human_development_index) & !is.na(continent) & continent %in% continents) %>%
  group_by(location) %>%
  summarize(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
            total_smokers = sum(female_smokers + male_smokers, na.rm = TRUE))

threshold <- median(severity$total_smokers)
severity$StatementTrue <- severity$total_smokers > threshold

# Calculate the percentage of continents where the statement is true
percentage <- sum(severity$StatementTrue) / nrow(severity) * 100

# Create the scatter plot
smoke <- ggplot(severity, aes(x = total_smokers, y = new_deaths)) +
  geom_point() +
  geom_hline(yintercept = threshold, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Relationship between Total Smokers and New Deaths from COVID-19",
       x = "Total Smokers",
       y = "New Deaths") +
  annotate("text", x = 5000, y = 1000, label = paste0("Total Countries: ", nrow(severity))) +
  annotate("text", x = 5000, y = 800, label = paste0("Statement True: ", round(percentage, 2), "%"),
            color = "blue")


#diabetes

diabetes <- covid_data %>%
  filter(!is.na(excess_mortality) & !is.na(human_development_index) & !is.na(continent) & continent %in% continents) %>%
  group_by(location) %>%
  summarize(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
            total_diabetes = sum(diabetes_prevalence, na.rm = TRUE))

threshold2 <- median(diabetes$total_diabetes)
diabetes$StatementTrue <- diabetes$total_diabetes > threshold2

# Calculate the percentage of continents where the statement is true
percentage2<- sum(severity$StatementTrue) / nrow(severity) * 100

# Create the scatter plot
diabetes_plot <- ggplot(diabetes, aes(x = total_diabetes, y = new_deaths)) +
  geom_point() +
  geom_hline(yintercept = threshold2, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Relationship between Diabetes and New Deaths from COVID-19",
       x = "Diabetes",
       y = "New Deaths") +
  annotate("text", x = 500, y = 750, label = paste0("Total Countries: ", nrow(diabetes))) +
  annotate("text", x = 500, y = 700, label = paste0("Statement True: ", round(percentage2, 2), "%"),
            color = "blue")


#function
library(ggplot2)
library(gridExtra)

plot_relationship <- function(covid_data) {
    severity <- covid_data %>%
        filter(!is.na(excess_mortality) & !is.na(human_development_index) & !is.na(continent) & continent %in% continents) %>%
        group_by(location) %>%
        summarize(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
                  total_smokers = sum(female_smokers + male_smokers, na.rm = TRUE))

    threshold <- median(severity$total_smokers)
    severity$StatementTrue <- severity$total_smokers > threshold

    # Calculate the percentage of continents where the statement is true
    percentage <- sum(severity$StatementTrue) / nrow(severity) * 100

    # Create the scatter plot
    smoke <- ggplot(severity, aes(x = total_smokers, y = new_deaths)) +
        geom_point() +
        geom_hline(yintercept = threshold, linetype = "dashed", color = "red") +
        theme_minimal() +
        labs(title = "Relationship between Total Smokers and New Deaths from COVID-19",
             x = "Total Smokers",
             y = "New Deaths") +
        annotate("text", x = 5000, y = 1000, label = paste0("Total Countries: ", nrow(severity))) +
        annotate("text", x = 5000, y = 800, label = paste0("Statement True: ", round(percentage, 2), "%"),
                 color = "blue")


    #diabetes

    diabetes <- covid_data %>%
        filter(!is.na(excess_mortality) & !is.na(human_development_index) & !is.na(continent) & continent %in% continents) %>%
        group_by(location) %>%
        summarize(new_deaths = sum(new_deaths_smoothed_per_million, na.rm = TRUE),
                  total_diabetes = sum(diabetes_prevalence, na.rm = TRUE))

    threshold2 <- median(diabetes$total_diabetes)
    diabetes$StatementTrue <- diabetes$total_diabetes > threshold2

    # Calculate the percentage of continents where the statement is true
    percentage2<- sum(diabetes$StatementTrue) / nrow(diabetes) * 100

    # Create the scatter plot
    diabetes_plot <- ggplot(diabetes, aes(x = total_diabetes, y = new_deaths)) +
        geom_point() +
        geom_hline(yintercept = threshold2, linetype = "dashed", color = "red") +
        theme_minimal() +
        labs(title = "Relationship between Diabetes and New Deaths from COVID-19",
             x = "Diabetes",
             y = "New Deaths") +
        annotate("text", x = 500, y = 750, label = paste0("Total Countries: ", nrow(diabetes))) +
        annotate("text", x = 500, y = 700, label = paste0("Statement True: ", round(percentage2, 2), "%"),
                 color = "blue")

    return(grid.arrange(smoke, diabetes_plot, nrow = 2))

}


#plot
plot_relationship(covid_data)




```
## Part C

Show how quickly different regions increased their hospitalization facilities, and whether
this led or lagged ICU admissions.

```{r}
library(ggplot2)
library(dplyr)


continents <- c("Asia", "Africa", "Europe", "North America", "South America", "Oceania")



days_since_first_infection <-  covid_data %>%
  filter(total_cases >= 1) %>%
  filter(continent %in% continents) %>%
  group_by(continent) %>%
  mutate(date = as.Date(date)) %>%
  arrange(continent) %>%
  mutate(DaysSinceCase1 = as.numeric(date - min(date))) %>%
  mutate(increase_facilities = hospital_beds_per_thousand,
         increase_icu_admissions = weekly_icu_admissions_per_million) %>% 
    dplyr::select(continent, date, increase_facilities,increase_icu_admissions, DaysSinceCase1) %>%
  ungroup()

complete_data <- na.omit(days_since_first_infection)

ggplot(data = complete_data) +
  geom_line(aes(x = DaysSinceCase1, y = increase_facilities, colour = continent), size = 1) +
  geom_line(aes(x = DaysSinceCase1, y = increase_icu_admissions, colour = continent), size = 1, linetype="dashed") +
  facet_wrap(~continent, nrow = 2) +
  labs(title = "Increase in Facilities and ICU Admissions over Time by Continent",
       x = "Days Since First Infection",
       y = "Count") +
  scale_color_discrete(name = "Continent") +
  theme_minimal()



library(scales)

ggplot(data = complete_data) +
  geom_bar(aes(x = DaysSinceCase1, y = rescale(increase_facilities), fill = "Facilities"), stat = "identity", width = 0.5) +
  geom_line(aes(x = DaysSinceCase1, y = rescale(increase_icu_admissions), colour = continent), size = 1.5) +
  facet_wrap(~continent, nrow = 2) +
  labs(title = "Increase in Facilities and ICU Admissions over Time by Continent",
       x = "Days Since First Infection",
       y = "Scaled Value") +
  scale_fill_manual(values = c("Facilities" = "#1f77b4"), guide = guide_legend(title = "Data")) +
  theme_minimal()




```


#Question 2: London weather
You mentioned you will do some analytics to prove to her the contrary - so you reached out to
your friend at the UK National Weather Service to get the data in Data/London to prove your
point.
Use any means possible to convince her of the state of the weather (and you may do some research
outside of what is provided, e.g. get other regions’ aggregates if you’d like to contextualize it)
- as she is a visual arts student she would naturally be more convinced by good looking figures
rich with relevant content.
You will also notice that there are two data files - one focussed on London, the other from the
CET CENTRAL ENGLAND, UK weather station (middle of the country). The latter is more
detailed, and column definitions are contained in the included


```{r}
weather <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/London/london_weather.csv")

# Load the necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)

# Assuming your dataset is stored in a variable called 'data'

# Convert the date variable to a Date format
weather$date <- as.Date(as.character(weather$date), format = "%Y%m%d")

# Calculate the average daily sunshine duration for each month
monthly_sunshine <- weather %>%
  group_by(year = year(date), month = month(date)) %>%
  summarize(avg_sunshine = mean(sunshine, na.rm = TRUE))

# Create a line plot showing the monthly variation in sunshine duration
ggplot(monthly_sunshine, aes(x = month, y = avg_sunshine, group = year)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(title = "Monthly Variation in Sunshine Duration",
       x = "Month", y = "Average Daily Sunshine Duration") +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  theme_bw()







# Load the necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)

# Convert the date variable to a Date format



# Create a dual-axis line plot
ggplot(weather, aes(x = date)) +
  geom_line(aes(y = max_temp, color = "Maximum Temperature"), size = 1) +
  geom_line(aes(y = min_temp, color = "Minimum Temperature"), size = 1, linetype = "dashed") +
  scale_color_manual(values = c("Maximum Temperature" = "red", "Minimum Temperature" = "blue")) +
  labs(title = "Temperature Extremes in London",
       x = "Date", y = "Temperature (°C)") +
  theme_bw() +
  theme(legend.position = "top")









```

```{r}
library(forecast)


```


##Question 3

##bands
Load data
```{r}
coldplay <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/Coldplay_vs_Metallica/Coldplay.csv")

# clean <- try[grepl(paste(c("Midnights \\(3am Edition\\)", "Taylor's Version", "deluxe version", "Deluxe Edition", "Live", "Radio", "Special", "Taylor Swift", "reputation", "Lover"), collapse = "|"), try$album, ignore.case = TRUE), ]


coldplay_studio <- coldplay[!grepl(paste(c("Live", "Tour Edition", "\\(Live\\)", "\\(Prospekts March Edition\\)", "Love in Tokyo"), collapse = "|"), coldplay$album_name, ignore.case = TRUE), ]


coldplay_studio <- coldplay_studio[!grepl(paste(c("Charlie Brown - Live from Glastonbury, 2011",
                   "Life Is for Living - Live from Glastonbury, 2011",
                   "Every Teardrop Is a Waterfall - Live from Glastonbury, 2011"), collapse = "|"), coldplay_studio$name, ignore.case = TRUE), ]
coldplay_studio <- coldplay_studio %>% rename(album=album_name)


metallica <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/Coldplay_vs_Metallica/metallica.csv")

metallica_studio <- metallica[!grepl(paste(c("Live",  "\\(Remastered 2021\\)", "\\(Remastered Deluxe Box Set\\)", "\\(Deluxe / Remastered\\)", "\\(Deluxe Box Set / Remastered\\)", "Live In "), collapse = "|"), metallica$album, ignore.case = TRUE), ]

rows_to_remove <- c("Live In Brazil (1993 \x93 2017)","Live In Argentina (1993 \x93 2017)", "Live In Chile (1993 \x93 2017)", "...And Justice For All","...And Justice for All (Remastered)", "Ride The Lightning (Remastered)","Kill Em All (Remastered)"  )
metallica_studio <- metallica_studio[!metallica_studio$album %in% rows_to_remove, ]
```

#popularity per album
```{r}

library(glue)
popularity_plot <- function(data, title) {
    data%>% 
  filter(!is.na(album)) %>%  # Exclude rows with NA album
  mutate(album = factor(album),  # Specify the correct order of the album names
         Text = ifelse(popularity == min(popularity, na.rm = TRUE), glue::glue("Lowest: {album}, ({name}) :{popularity}"), 
                       ifelse(popularity == max(popularity, na.rm = TRUE), glue::glue("Highest: {album}, ({name}) :{popularity}"),
                              NA_character_))) %>% 
  ggplot() + 
  geom_boxplot(aes(x = album, y = popularity, fill = album), alpha = 0.4) +
  geom_jitter(aes(x = album, y = popularity, color = album), size = 3, alpha = 0.8) + 
  ggrepel::geom_text_repel(aes(x = album, y = popularity, label = Text), force = TRUE, size = 3) +  # Adjust the size parameter for smaller text
  theme_classic() + 
  labs(x = "Albums", y = "Popularity",
       title = title) + 
  scale_color_hue(l = 40, c = 35) + 
  guides(fill = FALSE, color = FALSE)+
    theme(axis.title.x = element_text(face = "bold", colour = "black"),
          axis.text.x = element_text(face = "italic", angle = 90),
          axis.title.y = element_text(face = "bold", colour = "black"),
          text = element_text(size = 12, family = "Times"))


}

popularity_plot(coldplay_studio, "Popularity per Album (Coldplay)")
popularity_plot(metallica_studio, "Popularity per Album (Metallica")



#for corr matrix maybe I will use this, but I think it will be too much to also include
# cor_table <- data.frame(round(cor(coldplay_studio[,5:15], method = "pearson", use = "complete.obs"),4))
# #plot correlation matrix
# cor_table %>%
#   rownames_to_column() %>%
#   pivot_longer(-1) %>%
#   ggplot(aes(x = name, y = rowname, fill = value)) +
#   geom_tile() +
#   geom_text(aes(label = value, family="Times")) +
#   coord_equal() +
#   scale_fill_gradient2(low = "blue3", mid = "white", high = "red3") +
#   labs(x = NULL, y = NULL) +theme_classic()+theme(text=element_text(size=12, 
#                                                                     family="Times"))



```


Style evolution
```{r}
#I suspect that most of the time popularity is linked with dancebility

#i will maybe use
# coldplay_studio %>%
#   mutate(correlation = round(cor(danceability, acousticness, method = "spearman"), 4)) %>%
#   mutate(relationship = ifelse(correlation < 0.5 && correlation > 0, "Weak positive",
#                        ifelse(correlation >= 0.5 && correlation <= 1, "Strong positive",
#                               ifelse(correlation > -0.5 && correlation < 0, "Weak negative",
#                                      ifelse(correlation <= -0.5 && correlation >= -1, "Strong negative",
#                                             "No correlation"))))) %>% 
#   ggplot(aes(x=danceability, y=acousticness,)) +
#   geom_point() +
#   geom_label(aes(label = paste0("Correlation: ", correlation), x = 0.4, y = 1), 
#              size = 2, alpha = 0.8, fill = "red") +
#   geom_label(aes(label = paste0("Strength of relationship: ", relationship), x = 0.4, y = 1), 
#              size =2, alpha = 0.8, fill = "red") +
#   theme_bw() +
#   labs(title = "Relationship between Acousticness and Danceability of Songs",
#        x = "Danceability",
#        y = "Popularity")




# Modify the data preparation steps for TF-IDF
library(dplyr)
library(tidyr)
library(tidytext)
library(ggplot2)


create_style_evolution_plot <- function(data, title) {
    tfidf <- data %>%
  distinct(album, acousticness, danceability, energy, instrumentalness, liveness, loudness, speechiness, valence, tempo) %>%
  arrange(album)
    
  long_data <- tfidf %>%
    tidyr::pivot_longer(cols = -album, names_to = "variable", values_to = "value")
  
  ggplot(long_data, aes(x = album, y = value, fill = variable)) +
    geom_col(show.legend = FALSE) +
    labs(x = "Variable", y = "Value", title = title) +
    theme_minimal() +
    facet_wrap(~variable, scales = "free", ncol = 5) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  
}



create_style_evolution_plot(coldplay_studio, "Evolution of Coldplay's Style")

create_style_evolution_plot(metallica_studio, "Evolution of Metallica's Style")






```

Now look at emotions 
```{r}
#emotions of album
classify_album_emotion <- function(data,title) {
  # Define the threshold values
  threshold <- list(acousticness = 0.2952089,
                    danceability = 0.4178667,
                    energy = 0.5618405,
                    instrumentalness = 0.2120298,
                    tempo = 124.3586,
                    loudness = -9.406095,
                    speechiness = 0.04195873,
                    valence = 0.2333286)

   data %>%
    mutate(emotion = ifelse(acousticness > threshold$acousticness &
                              danceability < threshold$danceability |
                              energy < threshold$energy &
                              instrumentalness > threshold$instrumentalness |
                              tempo < threshold$tempo &
                              loudness > threshold$loudness |
                              speechiness < threshold$speechiness |
                              valence < threshold$valence, "Sad", "Happy")) %>%
    group_by(album, emotion) %>%
    summarise(count = n()) %>%
    mutate(proportion = count / sum(count)) %>% 
       ggplot(aes(x = "", y = proportion, fill = emotion)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar(theta = "y") +
    facet_wrap(~album) +
    labs(x = NULL, y = NULL, fill = "Emotion", title = title) +
    theme_minimal() +
    theme(legend.position = "right") +
    scale_fill_manual(values = c("Sad" = "lightblue", "Happy" = "gold"))
}



classify_album_emotion(coldplay_studio, "Emotion Classification by Album (Coldplay)")
classify_album_emotion(metallica_studio, "Emotion Classification by Album(Metallica)")


```
#Question 4

You’ve been supplied with data on most titles available on Netflix over time. Note
for the column definitions: IMDb (an abbreviation of Internet Movie Database) is an online
database of information related to films, television series, home videos, video games, and streaming
content online – including cast, production crew and personal biographies, plot summaries,
trivia, ratings, and fan and critical reviews.
Given that you’re a respected quantitative analyst, you’ve been given full freedom by your
superiors to explore the topic and supply insights from the data that you deem interesting.

load data
```{r}
credits <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/netflix/credits.csv")
titles <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/netflix/titles.csv")
netflix <- full_join(credits, titles, by="id")
```
# insights
Trends in Netflix Content:
```{r}
library(ggplot2)


# Calculate the count of each genre
genre_counts <- netflix %>%
  unnest_tokens(genre, genres) %>%
  count(genre, sort = TRUE)

# Get the most popular genre
most_popular_genre <- genre_counts$genre[1]

# Calculate the count of each type (movies or shows)
type_counts <- netflix %>%
  count(type)

# Get the most popular type
most_popular_type <- type_counts$type[which.max(type_counts$n)]

# Create a bar plot for genre counts


# Create a bar plot with different colors for each genre
ggplot(genre_counts, aes(x = genre, y = n, fill = genre)) +
  geom_bar(stat = "identity") +
  labs(x = "Genre", y = "Count", title = "Genre Distribution") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  annotate("text", x = most_popular_genre, y = genre_counts$n[1], label = paste("Most Popular Genre:", most_popular_genre),
           vjust = -0.8, hjust = 0, color = "red") +
  coord_flip() 


# Create a bar plot for type counts
ggplot(type_counts, aes(x = type, y = n)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(x = "Type", y = "Count", title = "Type Distribution") +
  geom_text(aes(label = n), vjust = -0.3, color = "black", size = 4) +
  annotate("text", x = most_popular_type, y = type_counts$n[which.max(type_counts$n)], label = paste("Most Popular Type:", most_popular_type),
           vjust = -1.5, hjust = 0, color = "red")


```

Country-wise Analysis (Choropleth Map):
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(maps)
library(ggthemes)

library(dplyr)
library(tidyr)
library(maps)
library(ggthemes)

# Load map data for the world
world_map <- map_data("world")

# Count the number of titles by country
country_counts <- netflix %>%
  unnest_tokens(country, production_countries) %>%
  count(country) %>%
  top_n(10, n) %>%
  arrange(desc(n))

# Define a named vector of country codes and names
countries <- c(
  "us" = "USA",
  "in" = "India",
  "gb" = "UK",
  "jp" = "Japan",
  "fr" = "France",
  "ca" = "Canada",
  "es" = "Spain",
  "kr" = "South Korea",
  "de" = "Germany",
  "cn" = "China"
)

# Replace country codes with country names
country_counts$country <- countries[country_counts$country]

# Calculate the percentage of movies made in each country
country_counts <- country_counts %>%
  mutate(percentage = n / sum(n) * 100)

# Merge data with map data
merged_data_map <- merge(world_map, country_counts, by.x = "region", by.y = "country", all.x = TRUE)

# Create a choropleth map with scaled fill color and percentage labels
ggplot(merged_data_map, aes(x = long, y = lat, group = group, fill = percentage)) +
  geom_polygon(color = "black", linewidth = 0.2) +
  labs(fill = "Percentage of Movies", title = "Country-wise Analysis") +
  theme_map()


 
 


```
Popular actor per genre and imdb score
```{r}

library(tidyverse)


# Define a named vector of country codes and names
top_countries <- c(
  "us" = "USA",
  "in" = "India",
  "gb" = "UK",
  "jp" = "Japan",
  "fr" = "France",
  "ca" = "Canada",
  "es" = "Spain",
  "kr" = "South Korea",
  "de" = "Germany",
  "cn" = "China"
)

# Find the top actor per country and genre based on IMDb score
filtered_data <- netflix %>%
  unnest_tokens(country, production_countries) %>%
  filter(country %in% names(top_countries)) %>%
  unnest_tokens(genre, genres) %>%
  group_by(country, genre) %>%
  filter(imdb_score == max(imdb_score, na.rm = TRUE)) %>%
  dplyr::select(country, genre, name, role,imdb_score)

if ("role" %in% colnames(filtered_data) && "ACTOR" %in% filtered_data$role) {
  top_actor <- filtered_data %>%
    filter(role == "ACTOR") %>%
    slice(1) %>%
    pull(name)

  top_actors <- filtered_data %>%
    filter(name %in% top_actor) %>%
    count(genre, name, country)

  ggplot(top_actors) +
     geom_point(aes(x = genre, y = count, size = imdb_score, colour = name))  +
    labs(x = "Genre", y = "Count", title = "Dominant Name per Genre") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(~ country, ncol = 2)
} else {
  print("No top actor found.")
}


filtered_data <- netflix %>%
  unnest_tokens(country, production_countries) %>%
  filter(country %in% names(top_countries)) %>%
  unnest_tokens(genre, genres) %>%
  group_by(country, genre) %>%
  filter(imdb_score == max(imdb_score, na.rm = TRUE)) %>%
  dplyr::select(country, genre, name, role, imdb_score)

table_actor <- filtered_data %>%
  filter(role == "ACTOR") %>%
  group_by(country) %>%
  summarise(dominating_actor = first(name), dominating_genre = first(genre)) %>%
  arrange(country)

table_director <- filtered_data %>%
  filter(role == "DIRECTOR") %>%
  group_by(country) %>%
  summarise(dominating_director = first(name), dominating_genre = first(genre)) %>%
  arrange(country)


knitr::kables(
  list(
    # the first kable() to change column names
    knitr::kable(
      table_actor, col.names = c('Country Code', 'Dominating Actor', 'Dominating Genre'), valign = 't'
    ),
    # the second kable() to set the digits option
    knitr::kable(table_actor, col.names = c('Country Code', 'Dominating Director', 'Dominating Genre'), valign = 't'
    ))
)



```



Movie recommendation example
```{r}



generate_movie_recommendation <- function(data, name, production_countries, imdb_score, genres) {
  # Filter movies based on the provided features
  filtered_movies <- data %>%
    filter(
      name %in% name,
      production_countries == production_countries,
      imdb_score >= imdb_score,
      imdb_score %in% imdb_score
    )

  # Sort the movies by popularity or score, depending on preference
  sorted_movies <- filtered_movies %>%
    arrange(desc(tmdb_popularity))

  # Return the top recommended movie
  if (nrow(sorted_movies) > 0) {
    top_movie <- sorted_movies[1, "title"]
    return(top_movie)
  } else {
    return("No movie recommendation found.")
  }
}

generate_movie_recommendation(netflix, name=="Caterina Murino", production_countries=="['US']", imdb_score>=8, genres=="['drama']")






```

# Question 5

You’ve been given a detailed database of app downloads from Google by your superior, head of
an App development company called AppallingDesigns.
You are looking to develop a new app, but need some background information into App design
trends.
Give your editor some insight into which categories are most profitable and also some insights
into app sizes, and more.
You were also given a dataset of Reviews in your folder. You can use this as well to supplement
your research - by considering some user feedback too


Load data
```{r}
apps <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/googleplay/googleplaystore.csv")
reviews <- read.csv("/Users/gabriellaneilon/Library/Mobile Documents/com~apple~CloudDocs/Masters/DatsciPractical23/data/googleplay/googleplaystore_user_reviews.csv")


apps_complete <- full_join(apps,reviews, by="App")


```

lets start coding
```{r}
apps_processed %>%
  mutate(total_earnings = Price * Installs)
# Calculate total earnings (Price * Installs) for each app
earnings <- apps_processed %>%
   mutate(total_earnings = Price * Installs)
 # Profitability by Category (Total Earnings)
 profitability <- earnings %>%
  group_by(Category) %>%
   summarize(total_earnings = sum(total_earnings, na.rm = TRUE))
 # Profitability by Category (Total Earnings)

apps_processed <- apps_complete %>%
  mutate(
    Price = as.numeric(gsub("\\$", "", Price)),
    Installs = as.numeric(gsub("[+,]", "", Installs)),
    Size = ifelse(grepl("M$", Size), as.numeric(gsub("M$", "", Size)) * 1000000, NA)
  ) %>%
  inner_join(
    earnings %>% 
      group_by(Category) %>% 
      summarize(total_earnings = sum(total_earnings, na.rm = TRUE)),
    by = "Category"
  ) %>%
  group_by(Category) %>%
  summarize(average_rating = mean(Rating, na.rm = TRUE), average_earnings = mean(total_earnings, na.rm = TRUE))

apps_processed <- na.omit(apps_processed)

library(ggplot2)


# Plot the earnings by category with rating
ggplot(apps_processed) +
  geom_segment(aes(x = Category, xend = Category, y = 0, yend = average_earnings), color = "gray") +
  geom_point(aes(x = Category, y = average_earnings, size = average_rating, colour = Category)) +
  labs(x = "Category", y = "Average Earnings", title = "Earnings by Category with Rating") +
  theme_minimal() +
  coord_flip()+
    scale_y_continuous(labels = function(y) format(y, scientific = FALSE))


```
Reviews
```{r}

#do the following: first look at how many postive and negative and neutral sentiments there are per category, then look at total reviews per category, then divideive the sentiments per category by the total reviews to get a percentage
library(tidyverse)
# Calculate the number of positive, negative, and neutral sentiments per category
sentiment_counts <- apps_complete %>%
  group_by(Category, Sentiment) %>%
  summarise(count = n()) %>% 
  filter(Sentiment != "nan" & !is.na(Sentiment))

#total sentiments 
total_sentiments <- sentiment_counts %>%
  group_by(Category) %>%
  summarise(total = sum(count))

# Calculate the percentage of each sentiment
sentiment_percentages <- sentiment_counts %>%
  left_join(total_sentiments, by = "Category") %>%
  mutate(percentage = count / total * 100)



# Plot the sentiment percentages as pie charts for each category
ggplot(sentiment_percentages, aes(x = "", y = percentage, fill = Sentiment)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(x = NULL, y = NULL, title = "Sentiment Percentages per Category") +
  scale_fill_manual(values = c("Positive" = "lightgreen", "Negative" = "pink", "Neutral" = "gold")) +
  theme_void() +
  facet_wrap(~ Category, ncol = 5) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_text(aes(label = paste0(round(percentage), "%")), position = position_stack(vjust = 0.5), size=4)



# Display all categories/rows in sentiment_percentages dataframe
knitr::kable(sentiment_percentages, col.names = gsub("[.]", " ", names(sentiment_percentages)), row.names = TRUE,caption = "An example table caption.")


```

